Comparing the Direction and Degree of Selection on Mutations Between SARS-CoV-2 Phylogenetic Clades

### 2022-11-21 expected_and_actual_counts_ratio/notebooks/expected_and_actual_counts_ratio.Rmd
Modify while test running notebook

```{r contingency_table}
# Iterate through rows and compare clade 1 observed and expected values to clade 2 observed and expected values.
#p_values <- data.frame(p_value=apply(comparisons, 1, FUN=function(c) {fisher.test(rbind(c(as.numeric(c["actual_count_1"]),as.numeric(c["rounded_expected_1"])),c(as.numeric(c["actual_count_2"]),as.numeric(c["rounded_expected_2"]))))$p.value}))

# Save and load the variable because calculating p-values may take a long time.
#save(p_values, file=paste("../results/pvalues", paste(clades, collapse="_"), option, ".Rda", sep=""))
load(paste("../results/pvalues", paste(clades, collapse="_"), option, ".Rda", sep=""))
```

@@

```{r volcano_all, echo=FALSE}
# Plot the uncorrected p-values.
ggplot(comparison_p_values, aes(x=fold_change, y=-log10(p_value)))+
  geom_point(alpha=0.5)+
  geom_hline(yintercept=-log10(0.05), col="red", linetype=2)+
  theme_bw()
```

@@

```{r volcano_corrected, echo=FALSE, fig.height=10}
# After dropping P = 1, plot the corrected p-values.
volcano <- comparison_p_values[which(comparison_p_values$p_value_corrected!=1),]
if("gene" %notin% colnames(volcano)){
  volcano$gene <- ifelse(volcano$gene_1==volcano$gene_2, volcano$gene_1, paste(volcano$gene_1, volcano$gene_2, sep="_"))
}
color <- colorRampPalette(brewer.pal(8,"Set1"))(length(unique(volcano$gene)))

>>>>

```{r contingency_table}
# Iterate through rows and compare clade 1 observed and expected values to clade 2 observed and expected values.
p_values <- data.frame(p_value=apply(comparisons, 1, FUN=function(c) {fisher.test(rbind(c(as.numeric(c["actual_count_1"]),as.numeric(c["rounded_expected_1"])),c(as.numeric(c["actual_count_2"]),as.numeric(c["rounded_expected_2"]))))$p.value}))

# Save and load the variable because calculating p-values may take a long time.
#save(p_values, file=paste("../results/pvalues", paste(clades, collapse="_"), option, ".Rda", sep=""))
#load(paste("../results/pvalues", paste(clades, collapse="_"), option, ".Rda", sep=""))
```

@@

```{r volcano_all, echo=FALSE}
# Plot the uncorrected p-values.
p <- ggplot(comparison_p_values, aes(x=fold_change, y=-log10(p_value)))+
  geom_point(alpha=0.5)+
  geom_hline(yintercept=-log10(0.05), col="red", linetype=2)+
  theme_bw()
```

@@

```{r volcano_corrected, echo=FALSE}
#fig.height=10?
# After dropping P = 1, plot the corrected p-values.
volcano <- comparison_p_values[which(comparison_p_values$p_value_corrected!=1),]
if("gene" %notin% colnames(volcano)){
  volcano$gene <- ifelse(volcano$gene_1==volcano$gene_2, volcano$gene_1, paste(volcano$gene_1, volcano$gene_2, sep="_"))
}
color <- colorRampPalette(brewer.pal(8,"Set1"))(length(unique(volcano$gene)))
